---
title: "pathrise"
output: html_notebook
---

This notebook analyzes the [raw data](https://www.pathrise.com/outcomes.pdf) from [Pathrise](https://www.pathrise.com/). Pathrise is a career mentoring startup that connects with students and early career individuals who are looking for a job. Sandesh Shrestha reached out to me on 12 Jan 2021 about my pursuit of a "data role in a big tech company." 

After taking a look at the website, I found self-reported outcomes of their clients. It must be stated that Pathrise does not only focus on clients who aspire to be in data science; they also help students who are seeking roles in Software Engineering, Product, Strategy, and Ops, Product Design, Marketing, and Sales.

#Load necessary libraries into workspace

```{r load necessary libraries}
pacman::p_load("tidyverse", "readxl", "tidytext", "wordcloud", "tm", "ggplot2")
```

#Load the data into the global environment

```{r load data}
# Load data into environment
df <- read_csv("data/numeric-data/pathrise_outcomes.csv")

```
#Preprocess dataframe

```{r clean dataframe}
#Remove dollar sign from salary column
df$Salary <- as.numeric(gsub('[$,]', '', df$Salary))
#Drop empty columns
df <- df %>% 
  select(-c("X4","X5","X6"))

# Change names and correct for data types
df <- df %>% 
  # Change names for column headers
  rename(salary_usd = Salary) %>% 
  rename(company = Company) %>% 
  rename(placement_months = `Placement (Time in Months)`) %>% 
  # Confirm data types for columns
  mutate(salary_usd = as.numeric(salary_usd)) %>% 
  mutate(company = as.character(company)) %>% 
  mutate(placement_months = as.numeric(placement_months))

# Preview dataframe
head(df, 10)

```

#Exploratory Data Analysis

##Mean values of salray and placement time
Descriptive statistics for the numeric values and wordcloud for the companies
```{r descriptive stats}

#Summarize stats for numeric data 
summary_df <- df %>% 
  summarise(mean_sal = mean(salary_usd), 
            mean_plac = mean(placement_months),
            med_sal = median(salary_usd),
            med_plac = median(placement_months),
            std_sal = sd(salary_usd),
            std_plac = sd(placement_months),
            min_sal = min(salary_usd),
            min_plac = min(placement_months),
            max_sal = max(salary_usd),
            max_plac = max(placement_months))

summary_df
```


```{r wordcloud}
# Visualize companies
corp <- SimpleCorpus(VectorSource(df$company))
tm <- corp %>% tm_map(function(x) removeWords(tolower(x), stopwords("en"))) %>%
  gsub('[[:punct:] ]+',' ', .) %>% 
  gsub('inc', ' ', .) %>% 
  gsub('unreported', ' ', .) %>% 
  gsub('data', ' ', .) %>% 
  gsub('com', ' ', .) %>% 
  gsub('llc', ' ', .)
# Generate wordcloud
wordcloud(tm, colors=brewer.pal(6,"Dark2"), random.order=FALSE)

```


#Distributions of salary and placement time
```{r distributions}
# Salary
df %>% 
  ggplot(aes(x = salary_usd))+
  geom_boxplot()+
  labs(x = "Salary (USD)")+
  scale_x_continuous(labels = scales::dollar_format())+
  theme(axis.text.y = element_blank())

# Placement Time
df %>% 
  ggplot(aes(x = placement_months))+
  geom_density()+
  labs(x = "Months Before Signing an Offer",
       y = "Density")

```

#Bar/Scatter Plot with placements and salary
```{r comparative plots}

df %>% 
  ggplot(aes(x = salary_usd))+
  geom_histogram(aes(y = ..density..),
                 bins = 30, 
                 breaks = seq(40000,300000, by = 15000),
                 col = "black",
                 alpha = 0.3) +
  geom_density(alpha = 0.2, fill = "#ff66a3") +
  labs(y = "Count",
       x = "Salary (USD)") +
  scale_x_continuous(labels = scales::dollar_format())

df %>% 
  ggplot(aes(x = placement_months))+
  geom_histogram(bins = 15,
                 col = "black",
                 alpha = 0.3) +
  labs(y = "Count",
       x = "Months Before Signing an Offer")

```
```{r comparative plots continued}
#Scatterplot
df %>%
  ggplot(aes(x= placement_months, y= salary_usd, label=company))+
  geom_point() +
  geom_text(aes(label=ifelse(company=="Microsoft",
                             as.character(company),'')), hjust=0,vjust=0)+
  labs(x = "Placement Time (months)", y = "Salary (USD)")

```

Examine jobs that were secured in four months or less
```{r short term to hire}

df %>%
  filter(placement_months <= 4) %>% 
  ggplot(aes(x = salary_usd))+
  geom_histogram(bins = 15, 
                 breaks = seq(40000,300000, by = 15000),
                 col = "black",
                 alpha = 0.3) +
  labs(title = "Salaries for Placements in Four Months or Less",
       y = "Count",
       x = "Salary (USD)") +
  scale_x_continuous(labels = scales::dollar_format())

df %>%
  filter(placement_months <= 4) %>%
  filter(salary_usd >= 80000 & salary_usd <= 130000) %>% 
  ggplot(aes(x = salary_usd))+
  geom_histogram(aes(y = ..density..),
                 col = "black",
                 alpha = 0.4,
                 bins = 15)+
  geom_density(alpha = 0.3, fill = "#ff66a3")+
  scale_x_continuous(labels = scales::dollar_format())
  
```

